Well-quasi-orderings on word languages

### 2025-09-30 src/infixes.tex
new: factor complexity uniformly recurrent morphic sequences

\AP
Let us conclude this section with a few remarks on the notion of \kl{factor
complexity} of languages. Recall that the \intro{factor complexity} of a
language $L \subseteq \Sigma^*$ is the function $f_L: \Nat \to \Nat$ such that
$f_L(n)$ is the number of distinct words of size $n$ in $L$. A language $L$ has
\intro{exponential factor complexity} if there exists a constant $C > 1$ such
that $f_L(n) \geq C^n$ for infinitely many $n \in \Nat$.

@@

It is the case that \kl{bounded languages} have at most a polynomial factor
complexity, and that the \kl{Thue-Morse sequence} has a linear factor
complexity. More generally, as we will see, most ``concrete'' examples of
infinite words have a low factor complexity: \kl{automatic sequences} have at
most a linear factor complexity, and \kl(morphic){morphic sequences} have at
most a polynomial factor complexity. In that sense, the
\cref{exponential-factor-complexity:lem} shows that one can still construct
``pathological'' examples of infinite words that are \kl{well-quasi-ordered}
for the \kl{infix relation}, but have a high factor complexity.

>>>>

\subsubsection{Factor Complexity}

\AP Let us conclude this section with a few remarks on the notion of \kl{factor
complexity} of languages. Recall that the \intro{factor complexity} of a
language $L \subseteq \Sigma^*$ is the function $f_L: \Nat \to \Nat$ such that
$f_L(n)$ is the number of distinct words of size $n$ in $L$. A language $L$ has
\intro{exponential factor complexity} if there exists a constant $C > 1$ such
that $f_L(n) \geq C^n$ for infinitely many $n \in \Nat$. A language $L$ has
\intro{subaffine factor complexity} if $f_L(n) \leq C \cdot n + K$ for some
constants $C, K > 0$ and all $n \in \Nat$.

@@

\todo[inline]{Rewrite this paragraph to cite precise results.}
It is the case that \kl{bounded languages} have at most a polynomial factor
complexity, and that the \kl{Thue-Morse sequence} has a linear factor
complexity. More generally, as we will see, most ``concrete'' examples of
infinite words have a low factor complexity: \kl{automatic sequences} have at
most a linear factor complexity, and \kl(morphic){morphic sequences} 
have a polynomial factor complexity. In
that sense, the \cref{exponential-factor-complexity:lem} shows that one can
still construct ``pathological'' examples of infinite words that are
\kl{well-quasi-ordered} for the \kl{infix relation}, but have a high factor
complexity.

Note that because of \cite[Theorem 24]{NIPR09}, any \kl{uniformly recurrent}
\kl{morphic sequence} has a linear factor complexity. In particular, we
conjecture that any \kl{ultimately uniformly recurrent} \kl{morphic sequence}
has a linear factor complexity too.
